Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2015-03-08 sas2015/content/background.tex
removing some old comments

\commentout{
To compute the meet-over-all-valid-paths solution, each node in the control-flow supergraph is paired with a \textit{fact} $d\in D\cup\{\mathbf0\},\,\mathbf0\notin D$, yielding the nodes $N^\#$ of the \textit{exploded supergraph} $G^\#=(N^\#,\,E^\#)$.
Roughly, for each node in the program, a fact denotes a binary property whose value (true or false) we want to find out.
The start node of the exploded supergraph is the node $(\startmain,\,\mathbf0)$.

The flow functions $F$ define the edges of the exploded supergraph.
Using the flow functions, the IFDS algorithm computes the interprocedurally \textit{realizable} paths from the start to the rest of the exploded graph's nodes. A realizable path is a valid path in the exploded supergraph that starts with the entry node $\startmain$.

If there is a realizable path from the node $(\startmain,\,\mathbf0)$ to a given node $(n,\,d),\,d\ne\mathbf0$, then the fact $d$ is considered to hold at node $n$. A path to a node $(n,\,\mathbf0)$ means that in the control-flow supergraph, there is a path from $\startmain$ to $n$.

In this way, the IFDS algorithm reduces the input data-flow problem to a graph-reachability problem.
\begin{example}
  In a taint analysis, $D$ is the set of variables in the program. If a fact $d\in D$ is reachable at a given node, then the variable is considered secret at that node. Otherwise, it is considered not secret. The question ``is $d$ secret at node $n$?'' becomes ``is there a realizable path from (\textsf{start}$_\texttt{main}$,\,\textbf{0}) to $(n,\,d)$?''.
\end{example}
\commentout{\begin{example} 
  In an available-expressions analysis, $D$ is the set of all expressions in the program. If an expression $d\in D$ is reachable at a certain node, it means that it does not need to be recomputed at that node. 
\end{example}}
}

%\mtodo{In Figure~\ref{fig:cc_edgefn_example}, should we say that $\psi$ corresponds to the function's return value?}
\begin{example}
  The exploded supergraph for Listing~\ref{list:ccexample} is shown in Figure~\ref{fig:cc_edgefn_example}.
  The labels on the edges will be explained in Section~\ref{sec:bgide}
  We can see that there is a realizable path, highlighted in bold,
  from the start node of the exploded graph to the variable \verb's' at~the~node~$\textsf{print(s)}$ in the \verb'B.bar' method. This means that \verb's' is considered secret at that node.

\commentout{  
    \begin{figure}%
      \hspace*{-30pt}
\scalebox{.9}{
      \tikzset{
        ashadow/.style={
          opacity=.25,
          shadow xshift=0.07,
          shadow yshift=-0.07
        },
      }
    \tikzstyle{supergraph}=[
      fill=greyblue,
      rounded corners,
      font=\small,
      align=right,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{supergraph_s}=[
      fill=bisque,
      rounded corners,
      font=\small,
      align=left,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{supergraph_f}=[
      fill=lightsalmonpink,
      rounded corners,
      font=\small,
      align=left,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{description}=[fill=white]
    \tikzstyle{n}=[fill=black,circle,inner sep=1.5pt]

\begin{tikzpicture}[>=stealth,on grid, auto]

% main method nodes
    \node [supergraph] (st_main) {\textsf{start}$_{\texttt{main}}$};
    \node [supergraph] (call_secret) [below = \dist of st_main.south east, anchor=east] {\textsf{call}$_\texttt{secret}$};
    \node [supergraph] (return_secret) [below = \dist of call_secret.south east,anchor=east] {\textsf{return}$_\texttt{secret}$\\\texttt{a = secret()}};
    \node [supergraph] (call_f_main) [below = \dist of return_secret.south east, anchor=east] {\textsf{call}$_\texttt{A.f}$};
    \node [supergraph] (return_f_main) [below = \dist of call_f_main.south east, anchor=east] {\textsf{return}$_\texttt{A.f}$\\\texttt{b = f(a)}};
    \node [supergraph] (end_main) [below = \dist of return_f_main.south east, anchor=east] {\textsf{end}$_{\texttt{main}}$};

    \node [n,label=$\mathbf{0}$] (st_main_z) [right=\ndd of st_main.east] {};
    \node [n,label=\texttt{a}] (st_main_x) [right=\ndd of st_main_z] {};
    \node [n,label=\texttt{b}] (st_main_y) [right=\ndd of st_main_x] {};

    \node [n] (call_secret_z) [right=\ndd of call_secret.east] {};
    \node [n] (call_secret_x) [right=\ndd of call_secret_z] {};
    \node [n] (call_secret_y) [right=\ndd of call_secret_x] {};

    \node [n] (return_secret_z) [right=\ndd of return_secret.east] {};
    \node [n] (return_secret_x) [right=\ndd of return_secret_z] {};
    \node [n] (return_secret_y) [right=\ndd of return_secret_x] {};

    \node [n] (call_f_main_z) [right=\ndd of call_f_main.east] {};
    \node [n] (call_f_main_x) [right=\ndd of call_f_main_z] {};
    \node [n] (call_f_main_y) [right=\ndd of call_f_main_x] {};

    \node [n] (return_f_main_z) [right=\ndd of return_f_main.east] {};
    \node [n] (return_f_main_x) [right=\ndd of return_f_main_z] {};
    \node [n] (return_f_main_y) [right=\ndd of return_f_main_x] {};

    \node [n] (end_main_z) [right=\ndd of end_main.east] {};
    \node [n] (end_main_x) [right=\ndd of end_main_z] {};
    \node [n] (end_main_y) [right=\ndd of end_main_x] {};
% main method edges
    \path[->](st_main_z) edge (call_secret_z);
    \path[->](call_secret_z) edge (return_secret_z);
    \path[->](return_secret_z) edge (call_f_main_z);
    \path[->](call_f_main_z) edge (return_f_main_z);
    \path[->](return_f_main_z) edge (end_main_z);  
    \path[->](return_f_main_y) edge (end_main_y);

%secret method nodes
    \node [supergraph_s] (st_secret) [above right = 2*\dist and 6*\dist of st_main] {\textsf{start}$_{\texttt{secret}}$};
    \node [supergraph_s] (return_secret_string) [below = \dist of st_secret.south west, anchor=west] {\texttt{return "secret"}};
    \node [supergraph_s] (end_secret) [below = \dist of return_secret_string.south west,anchor = west] {\textsf{end}$_\texttt{secret}$};

    \node [n,label=$\mathbf{0}$] (st_secret_z) [left=\ndd of st_secret.west] {};
    \node [n] (return_secret_string_z) [left=\ndd of return_secret_string.west] {};
    \node [n] (end_secret_z) [left=\ndd of end_secret.west] {};
%secret method edges
    \path[->](st_secret_z) edge (return_secret_string_z); 
    \path[->](return_secret_string_z) edge (end_secret_z); 
    \path[->,dashed](call_secret_z) edge [out=30,in=180] (st_secret_z);
    \path[->,dashed](end_secret_z) edge [out=240,in=40] (return_secret_z);
    \path[->,dashed](end_secret_z) edge [out=260,in=30] (return_secret_x);
    \path[->](return_secret_x) edge (call_f_main_x);
    \path[->](call_f_main_x) edge (return_f_main_x);
    \path[->](return_f_main_x) edge (end_main_x);

%f method nodes
    \node [supergraph_f] (st_f) [below=2*\dist of end_secret.south west,anchor=west] {\textsf{start}$_\texttt f$};
    \node [supergraph_f] (if) [below = \dist of st_f.south west,anchor=west] {\texttt{if (s == null)}};
    \node [supergraph_f] (return_x) [below right = \dist and 2.5*\dist of if.south] {\texttt{return s}};
    \node [supergraph_f] (call_f_f) [below=\dist of if.south west,anchor=west] {\textsf{call}$_\texttt{f}$};
    \node [supergraph_f] (return_f_f) [below = \dist of call_f_f.south west,anchor=west] {\textsf{return}$_\texttt{f}$\\\texttt{r = f("not secret")}};
    \node [supergraph_f] (return_r) [below = \dist of return_f_f.south west,anchor=west] {\texttt{return r}};
    \node [supergraph_f] (end_f) [below=\dist of return_r.south west,anchor=west] {\textsf{end}$_\texttt f$};

     \node [n,label=\texttt{s}] (st_f_s) [left=\ndd of st_f.west] {}; 
     \node [n,label=\texttt{r}] (st_f_r) [left=\ndd of st_f_s] {};
     \node [n,label=$\mathbf{0}$] (st_f_z) [left=\ndd of st_f_r] {};

     \node [n] (if_s) [left=\ndd of if.west] {};
     \node [n] (if_r) [left=\ndd of if_s] {};
     \node [n] (if_z) [left=\ndd of if_r] {};

     \node [n] (call_f_f_s) [left=\ndd of call_f_f.west] {};
     \node [n] (call_f_f_r) [left=\ndd of call_f_f_s] {};
     \node [n] (call_f_f_z) [left=\ndd of call_f_f_r] {};

     \node [n] (return_f_f_s) [left=\ndd of return_f_f.west] {};
     \node [n] (return_f_f_r) [left=\ndd of return_f_f_s] {};
     \node [n] (return_f_f_z) [left=\ndd of return_f_f_r] {};

     \node [n] (return_r_s) [left=\ndd of return_r.west] {};
     \node [n] (return_r_r) [left=\ndd of return_r_s] {};
     \node [n] (return_r_z) [left=\ndd of return_r_r] {};

     \node [n] (end_f_s) [left=\ndd of end_f.west] {};
     \node [n] (end_f_r) [left=\ndd of end_f_s] {};
     \node [n] (end_f_z) [left=\ndd of end_f_r] {};

     \node [n,label=$\mathbf0$] (return_x_z) [right=\ndd of return_x.east] {};
     \node [n,label=\texttt{r}] (return_x_r) [right=\ndd of return_x_z] {};
     \node [n,label=\texttt{s}] (return_x_s) [right=\ndd of return_x_r] {};
%f method edges
    \path[->](st_f_z) edge (if_z);
    \path[->](if_z) edge (call_f_f_z);
    \path[->](call_f_f_z) edge (return_r_z);  
    \path[->](return_r_z) edge (end_f_z);
    \path[->](if_z) edge [out=-43,in=127] (return_x_z);
    \path[->](return_x_z) edge [out=270,in=25] (end_f_z);
    \path[->](st_f_s) edge (if_s);
    \path[->](if_s) edge (call_f_f_s);
    \path[->](call_f_f_s) edge (return_r_s);  
    \path[->](return_r_s) edge (end_f_s);
    \path[->](if_s) edge [out=-30,in=110] (return_x_s);
    \path[->](return_x_s) edge [out=270,in=25] (end_f_s);
    \path[->](return_f_f_r) edge (return_r_r);
    \path[->](return_r_r) edge (end_f_r);
    \path[->,dashed](call_f_main_z) edge [out=20,in=160] (st_f_z);
    \path[->,dashed](end_f_z) edge [out=150,in=-20] (return_f_main_z);
    \path[->,dashed](call_f_f_z) edge [out=110,in=250] (st_f_z);
    \path[->,dashed](end_f_z) edge [out=110,in=250] (return_f_f_z);
    \path[->,dashed](call_f_main_x) edge [out=32,in=112] (st_f_s);
    \path[->,dashed](end_f_s) edge [out=155,in=-40] (return_f_main_y);
    \path[->,dashed](end_f_s) edge [out=120,in=-50] (return_f_f_r);
    \path[->,dashed](end_f_r) edge [out=110,in=250] (return_f_f_r);
    \path[->,dashed](end_f_r) edge [out=150,in=-50] (return_f_main_y);
\end{tikzpicture}
}
  \caption{The exploded supergraph corresponding to a taint analysis for the input program in Listing~\ref{list:examplejava}}%
  \label{fig:exampleexploded}%
  \end{figure}
}
\end{example}

\commentout{
To convert a representation relation $R_f$ back into the original flow function $f$, we can use the \textit{interpretation} function $\denote {R_f}$:
\begin{align}
  f&=\denote{R_f}\notag\\
  &=\lambda D_1\,.\,\left(\left\{d_2\ |\ \exists d_1\in D_1:\,(d_1,\,d_2)\in R_f\right\}\cup\{d_2\ |\ (\mathbf0,\,d_2)\in R_f\}\right)
  \setminus\{\mathbf0\}\,.
\end{align}
}

\commentout{
A practical implementation of the IFDS algorithm generally takes as input a representation of the
exploded supergraph edges $E^\sharp$ instead of explicit data-flow functions $M_F$. This is convenient because
the two are equivalent in expressiveness, and because the IFDS algorithm works internally with the exploded
supergraph. More specifically, the input generally provides a function
$f: N^* \times D \times N^* \to 2^D$. Given a (non-exploded) supergraph edge $n \to n'$ and a data-flow fact $d$,
$f(n,d,n')$ returns the set of all $d'$ such that the exploded supergraph contains the edge $(n,d) \to (n',d')$.
For convenience, the function $f$ can be split up into separate functions that handle the cases when
the $n\to n'$ is an intraprocedural edge, a call-to-start edge, or an end-to-return edge.
%\mtodo{Do we leave the start-to-return edge out as a special case of intraprocedural edge?}
}

%We presented an overview of the IFDS analysis. IFDS problems are transformed into IDE problems by the correlated-calls analysis. The IDE framework is described in the next section.

\subsection{IDE}\label{sec:bgide}
\commentout{
There exists an entire class of data-flow problems that cannot be formulated as IFDS problems. Informally, the problems cannot be formulated as decision problems. For instance, a constant-propagation problem asks, for each variable $v$ in the program, ``if $v$ is a constant at a given program point, what is $v$'s value?''.
The questions asked by constant propagation are of the form ``if property $X$ \textit{($v$ being a constant)} is true at program point~$Y$, what is the value of some property~$Z$ \textit{(the value of the constant)} corresponding to~$X$?''. It turns out that problems with such questions can often be solved by the IDE algorithm.

The IDE framework is an expressive extension to IFDS that was created by the same authors in 1996.
The problems that IDE can solve include, but are not limited to, IFDS problems~\cite{reps1995precise}. Just as the IFDS algorithm, the IDE algorithm is suitable for data-flow analyses that can be encoded with interprocedural, distributive flow functions. However, in IDE, the domain of the flow functions is not restricted to sets $D$ of data-flow facts. The IDE domain of a flow function consists of \textit{environments} that map data-flow facts from the set $D$ to lattice elements.

As an example, in a constant propagation problem, an IDE environment would map each variable to the (possibly) constant value that it is bound to. To illustrate the distinction between IFDS and IDE we could say that IFDS can find out which variables in a program are constants, whereas IDE can additionally retrieve the values of the constant variables.
Instead of just telling us whether a fact holds or not, the IDE analysis can provide us with additional information about facts.

Just as in the IFDS analysis, the IDE algorithm reduces a data-flow problem to a graph-reachability problem. Additionally, for each program point, the algorithm computes an \textit{environment} $\textsf{Env}(D,\,L)$, where data-flow facts are mapped to values of a lattice $L$.
}

@@

\commentout{
\begin{mdelete}
For example, using the IDE analysis, we can encode a restricted version of a constant-propagation analysis\footnote{%
  In the general case, constant propagation cannot be encoded with distributive flow functions and is therefore not suitable for an IDE analysis~\cite{muller2001complexity}.
}. The data-flow facts  correspond to program variables, and the lattice incorporates all possible values for constants. 
If a fact $d$ in the exploded supergraph is reachable at node $n$, and $\textsf{Env}(d)\notin\{\bot,\,\top\}$, it means that the variable associated with $d$ is a constant. Furthermore, the value of the constant can be inferred from the environment for the corresponding node and is equal to $\textsf{Env}(d)$.
\end{mdelete}
}

@@

\commentout{
and
\begin{equation}
  \Omega=\lambda d\,.\,\top
\end{equation}
is the top element in the environment lattice~$\mathsf{Env}(D,\,L)$.
}

\commentout{
The IDE analysis is a generalization of the IFDS analysis: every IFDS problem can be converted into an equivalent IDE problem~\cite{reps1995precise}. The equivalent problem can be solved by the IDE algorithm, and the result converted into an IFDS result. In~an~IFDS-equivalent IDE problem, the~graph~$G^*$ and the~set~$D$ of~data-flow facts remain the~same. The $L$ lattice is a two-point lattice: if a fact is mapped to the top (bottom) element, then it is reachable (unreachable). The conversion between IFDS and IDE problems is discussed in detail in Section~\ref{seq:transIfdsIde}.
}

\commentout{
\subsubsection{Environment Transformers}
For each node in the CFG, the result of an IDE analysis computes an environment $\textsf{Env}(D,\,L)$, which is a map from data-flow facts to lattice elements.

Instead of flow functions that show how to propagate facts, the IDE framework uses distributive environment transformers to propagate environments. For each edge $(n_1,\,n_2)$ in the control-flow supergraph, an environment transformer indicates how the environment at node $n_1$ is modified at node $n_2$.

From Section~\ref{sec:overviewifds} we know that flow functions can be represented with exploded-graph edges. To represent environment transformers, we will construct \textit{labeled} exploded-graph edges, where each edge is associated with a distributive \textit{micro function}%
\footnote{See~Sagiv et al.~\cite{sagiv1996precise} for a formal definition of the representation relation for environment transformers.}
$f:\,L\to L$. A~micro function shows how to change a lattice element for a given node and fact.

For instance, if an IDE problem is equivalent to an IFDS problem, the edges of the exploded supergraph are the same for both problems. In the IDE problem, the edges of the exploded supergraph are labeled with identity micro functions.

We extend the meet operator to work on micro functions by defining
\begin{equation}
  (f_1\sqcap f_2)(l)=f_1(l)\sqcap f_2(l)
\end{equation}
for all $l\in L$.
}

%In IDE problems, the auxiliary fact analogous to $\mathbf0$ in IFDS is denoted as $\Lambda$.

\commentout{
\begin{mdelete}
\begin{example}\label{ex:constprop}
  One version of the constant propagation analysis that can be encoded with IDE is \textit{linear constant propagation}. A linear constant propagation analysis can detect constants of the form $a\cdot x+b$, where $a$ and $b$ are integers and $x$ is a variable. In particular, a variable can only be considered constant if it depends on at most one other constant variable: even if $y$ and $z$ are variables that are considered constant, the variable $x=y+z$ will be considered not constant. If we encoded the analysis in a way to handle non-linear constant assignments, we would have to use non-distributive flow functions, which would violate the requirements of the IDE algorithm.

  For linear constant propagation, the $L$ lattice consists of the set of integers $\mathbb Z$, a top element denoting ``not a constant'', and a bottom element denoting an unknown value.
  The meet of two lattice elements is defined as follows: for any lattice element $l\in L$,
  \[
    \top\sqcap l=\top\qquad\text{and}\qquad\bot\sqcap l=l.
  \]
  For two lattice elements $l_1,\,l_2\in\mathbb Z$, $$l_1\sqcap l_2=\top.$$

  We define the addition and multiplication operations on lattice elements $l\in L$ and integers $c\in\mathbb Z$ as follows:
  \[
    l+c=\begin{cases}
      \bot&\text{if }l=\bot;\\
      \top&\text{if }l=\top;\\
      l+c&\text{otherwise.}
    \end{cases}
    \qquad
    c\cdot l=\begin{cases}
      \bot&\text{if }l=\bot;\\
      \top&\text{if }l=\top;\\
      c\cdot l&\text{otherwise.}
    \end{cases}
  \]
   Let the function $M$ that maps supergraph edges to environment transformers be defined in the following way:
  \[
    M=\lambda((n_1,\,n_2))\,.\,
    \begin{cases}
      \lambda\textsf{env}\,.\,\textsf{env}[x\to a\cdot\textsf{env}(y)+c\,]
        &\text{if $n_1$ contains an assignment}\\&\text{$x=a\cdot y+c$, where $y$ is a variable}\\&\text{and $a,\,c$ are constants;}\\
      \id
        &\text{otherwise.}
    \end{cases}
  \]
  Here, we denote with $\textsf{env}[x\to a]$ an environment \textsf{env} in which the key $x$ is mapped to $a$, and all other keys $y\ne x$ are mapped to their old values $\textsf{env}(y)$.
  When $M$ is applied to an edge whose source node contains an assignment for a variable $x$, $M$ returns an environment transformer that updates the argument environment with a new value for $x$.

  Consider the following program:

\inputMinted{java}{constprop.java}
  For the edges $e_1,\,e_2,\,e_3$, and $e_4$ that start at the first, second, third, and fourth instruction, $M$ creates the following environment transformers:
  \begin{align*}
    M(e_1)
    &=\lambda\textsf{env}\,.\,\textsf{env}[\texttt u\to 1]\\
    M(e_2)
    &=\lambda\textsf{env}\,.\,\textsf{env}[\texttt v\to\textsf{env}(\texttt u)+2]\\
    M(e_3)
    &=\lambda\textsf{env}\,.\,\textsf{env}[\texttt w\to\top]\\
    M(e_4)
    &=\lambda\textsf{env}\,.\,\textsf{env}[\texttt u\to 5].
  \end{align*}

  The corresponding labeled exploded supergraph is shown in Figure~\ref{fig:constprop}.

  The result of the analysis yields a map from nodes to environments. Each environment maps variables to elements of the constant-propagation lattice. The environment at the last node will look as follows:
  \[
    \{(\texttt u,\,5),\ (\texttt v,\,3),\ (\texttt w,\,\top)\}.
  \]
\end{example}

\begin{figure}
  \centering
    \tikzstyle{n}=[fill=black,circle,inner sep=1.5pt,sm]
    \tikzstyle{sm}=[font=\footnotesize]
\begin{tikzpicture}[>=stealth,on grid, auto]
    \node [n,label=$\mathbf0$] (zero){};
    \node [n] (zero_) [below=\edge of zero] {};
    \node [n,label={\texttt u}] (u) [right=\dist of zero] {};
    \node [n] (u_) [below=\edge of u] {};
    \node [n,label={\texttt v}] (v) [right=\dist of u] {};
    \node [n] (v_) [below=\edge of v] {};
    \node [n,label=\texttt{w}] (w) [right=\dist of v] {};
    \node [n] (w_) [below=\edge of w] {};
    \node [n] (zero__) [below=\edge of zero_]{};
    \node [n] (u__) [below=\edge of u_]{};
    \node [n] (v__) [below=\edge of v_]{};
    \node [n] (w__) [below=\edge of w_]{};
    \node [n] (zero___) [below=\edge of zero__]{};
    \node [n] (u___) [below=\edge of u__]{};
    \node [n] (v___) [below=\edge of v__)]{};
    \node [n] (w___) [below=\edge of w__]{};  
    \node [n] (zero____) [below=\edge of zero___]{};
    \node [n] (u____) [below=\edge of u___]{};
    \node [n] (v____) [below=\edge of v___)]{};
    \node [n] (w____) [below=\edge of w___]{};

    \path[->](zero) edge node[left,sm]{$\id$} (zero_);
    \path[->](zero) edge node[right,sm]{$\lambda l.1$} (u_);
    \path[->](zero_) edge node[left,sm]{$\id$} (zero__);
    \path[->](u_) edge node[right,sm]{$\lambda l.l+2$} (v__);
    \path[->](u_) edge node[left,sm]{$\id$} (u__);
    \path[->](zero__) edge node[left,sm]{$\id$} (zero___);
    \path[->](u__) edge node[left,sm]{$\id$} (u___);
    \path[->](v__) edge node[left,sm]{$\id$} (v___);

    \path[->,dashed,color=gray](v) edge node[left,sm]{$\id$} (v_);
    \path[->,dashed,color=gray](w) edge node[left,sm]{$\id$} (w_);
    \path[->,dashed,color=gray](w_) edge node[right,sm]{$\id$} (w__);

    \path[->] (zero___) edge node[left,sm]{$\id$} (zero____);
    \path[->] (u___) edge node[left,sm]{$\lambda l.5$} (u____);
    \path[->] (v___) edge node[left,sm]{$\id$} (v____);
    \path[->,dashed,color=gray] (w___) edge node[left,sm]{$\id$} (w____);
  \end{tikzpicture}
  \caption[A labeled exploded supergraph for a constant-propagation analysis described in Example~\ref{ex:constprop}]{A labeled exploded supergraph for a constant-propagation analysis described in Example~\ref{ex:constprop}. The dashed edges are edges not reachable from the entry node.}%
  \label{fig:constprop}%
  \end{figure}
\end{mdelete}
}

\commentout{
In this way, each edge in the exploded graph is labeled with a micro function. The mapping from exploded-graph edges to the corresponding micro functions is stored in \textit{edge functions},  denoted as \textsf{EdgeFn}$:\,E^\#\to(L\to L)$.
}

\subsubsection{Overview of the IDE Algorithm}\label{sec:ideoverview}
Just as any distributive function from $2^D$ to $2^D$ can be represented with a
representation relation, it is also possible to represent any distributive functions from
$\textsf{Env}(D,L)$
to
$\textsf{Env}(D,L)$
with a \textit{pointwise representation}. A pointwise representation is a bipartite graph
with the same nodes~\footnote{The IDE literature uses the symbol $\Lambda$ for the node that
    is denoted $\mathbf0$ in the IFDS literature. We use $\mathbf0$ throughout this paper for consistency.}
and edges as a representation relation, except that each edge is labelled
with a \textit{micro-function}, which is a function from $L$ to $L$. 
\commentout{
The pointwise representation of an environment transformer 
$t : \textsf{Env}(D,L) \to \textsf{Env}(D,L)$
is defined as
\newcommand{\microedge}[3]{#1\xrightarrow{#3}#2}
\begin{align*}
    R_t =& \{ \microedge{\mathbf0}{\mathbf0}{\lambda l.l} \} \cup
    \{\microedge{\mathbf0}{d_j}{\lambda l.t(\top_{\textsf{Env}})(d_j)} \mid d_j \in D\}\\
    &\cup
    \{\microedge{d_i}{d_j}{\lambda l.t(\top_{\textsf{Env}}[d_i\to l])(d_j)} \mid d_i, d_j \in D\}.
\end{align*}
}
%Let $\top_{\textsf{Env}} = \lambda d.\top$ be the environment that maps every element of $D$ to $\top$.
%\ftodo{
   %What is the relationship between $\Omega$ and $\Delta$?
%}
%\otodo{What do you mean by $\Delta$?}
Thanks to distributivity, every environment transformer 
$t : \textsf{Env}(D,L) \to \textsf{Env}(D,L)$
can be decomposed into its effect on $\top_{\textsf{Env}}$ and on a set of environments $\top_{\textsf{Env}}[d_i\to l]$
that map every element to $\top$ except one ($d_i$):
\[
    t(m)(d_j) = \lambda l. t(\top_{\textsf{Env}})(d_j) \sqcap \bigsqcap_{d_i\in D} \lambda l. t(\top_{\textsf{Env}}[d_i\to l])(d_j).
\]
The functions $\lambda l. \cdots$ in this decomposition are the micro-functions that
appear on the edges of the pointwise representation edges from $\mathbf0$ to each $d_j$ and from
each $d_i$ to each $d_j$.%
\footnote{The IDE paper defines a more complicated but equivalent set of micro-functions
that eliminate some duplication of computation.}
The absence of an edge in the pointwise representation from some $d_i$ to some $d_j$ is
equivalent to an edge with micro-function $\lambda l.\top$.

@@

%Similar to the IFDS algorithm,
A practical implementation of the IDE algorithm requires
the input data-flow functions to be provided in their pointwise representation
as exploded supergraph edges labelled with micro-functions.
%\mtodo{Should we cite the ``Practical extensions'' paper here that talks about the on-demand IFDS algorithm?}
Specifically,
the input is generally provided as a function $\textsf{EdgeFn}: (N^* \times D) \times (N^* \times D) \to F$,
where $F$ is the set of representations of micro-functions from $L$ to $L$.
Given an exploded supergraph edge $e=(n,\,d) \to (n',\,d')$,
$\textsf{EdgeFn}((n,\,d),\,(n',\,d'))$ returns the micro-function that appears on the exploded supergraph edge $e$.
%Like in the IFDS algorithm,
In an implementation, it can be convenient to split the function $\textsf{EdgeFn}$ into separate functions that handle the cases when $n\to n'$ is an intraprocedural edge, a call-to-return edge, a call-to-start edge, or an end-to-return edge.

\commentout{
Given a labeled exploded supergraph, the IDE algorithm computes the environments for all nodes in the CFG.

The algorithm first computes the lattice elements $l_{n,\,d}$ that correspond to each reachable node $(n,\,d)$ in the exploded supergraph. The union of the exploded nodes $(n,\,d)$ for a given control-flow node $n$, mapped to the corresponding lattice elements $l_{n,\,d}$, form the environment $\textsf{Env}_n$  for that node:
\begin{equation}
  \textsf{Env}_n=\left\{(d,\,l_{n,\,d})\ |\ (n,\,d)\in N^\#\right\}\,.
\end{equation}

The overall idea behind computing the lattice elements $l_{n,\,d}$ is the following. For each interprocedurally realizable path
\[
  p=\left[(\startmain,\,\mathbf0),\,(n_1,\,d_1),\,\dots,\,(n_k,\,d_k)\right]
\]
 that starts with the entrypoint of the exploded supergraph, we compute the micro function $f_p$ that corresponds to $p$. The micro function consists of the composition of all individual micro functions with which the edges of $p$ are labeled:
\begin{align}
  f_p=\textsf{EdgeFn}((n_{k-1},\,d_{k-1}),\,(n_k,\,d_k))\,\circ\ldots\circ
      \textsf{EdgeFn}(\startmain,\,\mathbf0),\,(n_1,\,d_1))\,.
\end{align}

Let the lattice element that $(n,\,d)$ is mapped to according to path $p$ be denoted as $l_{n,\,d}^p$. As shown in~Sagiv et al.~\cite{sagiv1996precise}, the lattice element can be obtained by applying $f_p$ to the bottom element:
\begin{equation}
 l^p_{n,\,d}=f_p(\bot)\,.
\end{equation}

Let $Q$ be the set of paths that start at the entry point and end at the given node $(n,\,d)$.
The lattice element $l_{n,\,d}$ is the meet of the lattice elements corresponding to all the paths in $Q$:
\[
  l_{n,\,d}=\bigsqcap_{q\in Q}l^q_{n,\,d}\,.
\]

This is a general outline of the IDE analysis. We use the IDE framework to improve the precision of IFDS problems in the presence of correlated calls.
The next section describes how this is done.
}

%\ftodo{Would it make sense to discuss an example of an IDE problem here?  Perhaps we can do this
%  without using much space by referring to an existing figure elsewhere in the paper?}
%  \otodo{See Example 4 on the previous page.}
%  \otodo{Now done in Example 5.}

>>>>

\begin{example}
  The exploded supergraph for Listing~\ref{list:ccexample} is shown in Figure~\ref{fig:cc_edgefn_example}.
  The labels on the edges will be explained in Section~\ref{sec:bgide}
  We can see that there is a realizable path, highlighted in bold,
  from the start node of the exploded graph to the variable \verb's' at~the~node~$\textsf{print(s)}$ in the \verb'B.bar' method. This means that \verb's' is considered secret at that node.
\end{example}

\subsection{IDE}\label{sec:bgide}

@@

@@

\subsubsection{Overview of the IDE Algorithm}\label{sec:ideoverview}
Just as any distributive function from $2^D$ to $2^D$ can be represented with a
representation relation, it is also possible to represent any distributive functions from
$\textsf{Env}(D,L)$
to
$\textsf{Env}(D,L)$
with a \textit{pointwise representation}. A pointwise representation is a bipartite graph
with the same nodes~\footnote{The IDE literature uses the symbol $\Lambda$ for the node that
    is denoted $\mathbf0$ in the IFDS literature. We use $\mathbf0$ throughout this paper for consistency.}
and edges as a representation relation, except that each edge is labelled
with a \textit{micro-function}, which is a function from $L$ to $L$.

Thanks to distributivity, every environment transformer 
$t : \textsf{Env}(D,L) \to \textsf{Env}(D,L)$
can be decomposed into its effect on $\top_{\textsf{Env}}$ and on a set of environments $\top_{\textsf{Env}}[d_i\to l]$
that map every element to $\top$ except one ($d_i$):
\[
    t(m)(d_j) = \lambda l. t(\top_{\textsf{Env}})(d_j) \sqcap \bigsqcap_{d_i\in D} \lambda l. t(\top_{\textsf{Env}}[d_i\to l])(d_j).
\]
The functions $\lambda l. \cdots$ in this decomposition are the micro-functions that
appear on the edges of the pointwise representation edges from $\mathbf0$ to each $d_j$ and from
each $d_i$ to each $d_j$.%
\footnote{The IDE paper defines a more complicated but equivalent set of micro-functions
that eliminate some duplication of computation.}
The absence of an edge in the pointwise representation from some $d_i$ to some $d_j$ is
equivalent to an edge with micro-function $\lambda l.\top$.

@@

A practical implementation of the IDE algorithm requires
the input data-flow functions to be provided in their pointwise representation
as exploded supergraph edges labelled with micro-functions.
%\mtodo{Should we cite the ``Practical extensions'' paper here that talks about the on-demand IFDS algorithm?}
Specifically,
the input is generally provided as a function $\textsf{EdgeFn}: (N^* \times D) \times (N^* \times D) \to F$,
where $F$ is the set of representations of micro-functions from $L$ to $L$.
Given an exploded supergraph edge $e=(n,\,d) \to (n',\,d')$,
$\textsf{EdgeFn}((n,\,d),\,(n',\,d'))$ returns the micro-function that appears on the exploded supergraph edge $e$.
%Like in the IFDS algorithm,
In an implementation, it can be convenient to split the function $\textsf{EdgeFn}$ into separate functions that handle the cases when $n\to n'$ is an intraprocedural edge, a call-to-return edge, a call-to-start edge, or an end-to-return edge.
